---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

<<<<<<< HEAD

## Loading and preprocessing the data



## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
=======
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(ggplot2)
library(dplyr)
library(data.table)
library(tidyr)
library(ggpubr)
library(stringr)
```

```{r ImportData, cache=TRUE}
## Loading and preprocessing the data
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", 
              destfile = "./assignment1data.zip")
unzip("./assignment1data.zip", exdir = "./Assignment1Data")
data <- read.csv("./Assignment1Data/activity.csv")
data$steps <- as.numeric(data$steps)
```

```{r PerDaySteps, fig.width=3.5, fig.height=1.8}
## What is mean total number of steps taken per day?
summary_steps <- data %>% group_by(date) %>% 
      summarise(Total_Steps_Per_day=sum(steps, na.rm = TRUE))

th <- theme(axis.text=element_text(size=12), axis.title = element_text(size =12), legend.text = element_text(size =12), strip.text.x = element_text(size=12))

ggplot(summary_steps, aes(x=Total_Steps_Per_day)) + geom_histogram(binwidth = 1000, color="black", fill="grey") + theme_bw() + th

mean <- mean(summary_steps$Total_Steps_Per_day)
median <- median(summary_steps$Total_Steps_Per_day)
```

The mean total number of steps taken per day is `r mean`. The median total number of steps taken per day is `r median`.

```{r ActivityPatter, fig.width=3.5,fig.height=1.8}
## What is the average daily activity pattern?
summary_pattern <- data %>% group_by(interval) %>% 
      summarise(Mean_Step_Each_Interval = mean(steps, na.rm = TRUE))
ggplot(summary_pattern, aes(x=interval, y=Mean_Step_Each_Interval)) + theme_bw() + th + geom_line()

max_steps <- summary_pattern[which.max(summary_pattern$Mean_Step_Each_Interval),]
```

At `r max_steps$interval` minute, on average across all the days, contains the maximum number of steps.

```{r NAvalues, fig.width=3.5,fig.height=1.8}
## Imputing missing values
num_na <- colSums(is.na(data))

#Find NAs and fill in with the mean number of steps in that interval of all days
row_na <- as.data.frame(which(is.na(data$steps)))
colnames(row_na) <- "Row_Num"
data2 <- data
for(i in 1:nrow(row_na)) {
      subset <- data2 %>% filter(interval == interval[row_na$Row_Num[i]])
      data2$steps[row_na$Row_Num[i]] <- mean(subset$steps, na.rm = TRUE)
}

#plot histogram
summary_steps2 <- data2 %>% group_by(date) %>% 
      summarise(Total_Steps_Per_day2=sum(steps, na.rm = TRUE))

th <- theme(axis.text=element_text(size=12), axis.title = element_text(size =12), legend.text = element_text(size =12), strip.text.x = element_text(size=12))

ggplot(summary_steps2, aes(x=Total_Steps_Per_day2)) + geom_histogram(binwidth = 1000, color="black", fill="grey") + theme_bw() + th

mean2 <- mean(summary_steps2$Total_Steps_Per_day2)
median2 <- median(summary_steps2$Total_Steps_Per_day2)
```

The total number of missing values in the dataset is `r num_na[1]`.

Using the dataset with NA values filled in, the mean total number of steps taken per day is `r mean2`. The median total number of steps taken per day is `r median2`. The total daily number of steps increased after filling in NA values.

```{r weekday versus weekend}
## Are there differences in activity patterns between weekdays and weekends?
data2$Day = weekdays(as.Date(data$date, "%Y-%m-%d"))
data2$Weekday_weekend = ifelse(data2$Day %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
avg_step <- data2 %>% group_by(Weekday_weekend, interval) %>% 
      summarise(avg = mean(steps))
ggplot(avg_step, aes(x = interval, y = avg)) + geom_line() + theme_bw() + th + 
      facet_grid(Weekday_weekend~., scales = "free") + ylab("Average Number of Steps")
```
>>>>>>> add three files
